---
title: "R Notebook"
output: html_notebook
---

Initiate setwd for all chunks
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/Uni/Doctorate/Ch2 Stressor/")
```

# Meta data
```{r Meta data}
# Libraries
library(tidyverse); library(readxl); library(patchwork); library(lme4); library(lmerTest); library(ozmaps); library(tidyterra); library(terra)


# Meta data
IID_block <- read_xlsx(path="FID Selection & ID & Randomisation_vers1.xlsx", sheet = 'real_IID_Block_Cell_12tray')
IID_block_sub <- IID_block %>% 
  select(IID, Mat_Line, Drought_sev, Drought_len, block_cell1, Status)

IID_block_sub$block_cell1 <- as.character(IID_block_sub$block_cell1)

# Maternal line data
parent_meta <- read.csv ("parent_meta.csv")
parent_meta_lib <- parent_meta %>% 
  dplyr::select(NSWID, sample_lib_NSW)

### Adding in meanCOI each maternal line
mquin_meta <- read.csv(file="~/RBGSyd_Technical Officer/MQuin/Processing Meta/Mquin_samples_pheno.csv")
Mat_meta <- mquin_meta %>% group_by(FID) %>% summarise(mean_COI=mean(COI, na.rm=T))
colnames(Mat_meta) <- c("Mat_Line", "mean_COI")

# Height-weight data
height_weight_data <- read.csv(file = "Height_weight_track.csv")
height_weight_data$Date <- as.Date(height_weight_data$Date, format = "%d/%m/%Y")
height_weight_data$Height..mm. <- as.numeric(height_weight_data$Height..mm.)
height_weight_data <- left_join(height_weight_data, IID_block)

height_weight_data_rmna <- height_weight_data %>% 
  filter(!is.na(Height..mm.))
```

```{r Maternal line sources}
Exp_Mat_Line <- unique(IID_block$Mat_Line)

parent_meta_rel <- parent_meta %>% filter (NSWID %in% Exp_Mat_Line); colnames(parent_meta_rel)[2] <- "Mat_Line"
parent_meta_rel <- left_join(parent_meta_rel, Mat_meta)

# worldclim_country("Australia", var="prec", path="~/Delete", res=0.5)
mnth_avg_prec <- raster::stack("~/Uni/Doctorate/Ch Planting/data/AUS_wc2.1_30s_prec.tif")

clim_res=NULL
for (i in 1:nrow(parent_meta_rel)){
  site=parent_meta_rel[i, "Pop"]
  lat=as.numeric(parent_meta_rel[i, "latitude"]); long=as.numeric(parent_meta_rel[i, "longitude"]) 
  prec = sum(raster::extract(mnth_avg_prec, cbind(long, lat)))
  clim_res <- as.data.frame(rbind(clim_res, cbind(Pop=site, longitude=as.numeric(long), latitude=as.numeric(lat), prec=prec)))
}
remove(mnth_avg_prec)

clim_res$longitude <- as.numeric(clim_res$longitude); clim_res$latitude <- as.numeric(clim_res$latitude)

parent_meta_rel <- unique(left_join(parent_meta_rel, clim_res)); parent_meta_rel$prec <- as.numeric(parent_meta_rel$prec)
IID_block <- left_join(IID_block, parent_meta_rel %>% dplyr::select(Mat_Line, mean_COI, longitude, latitude))
###
```

```{r}
### Plotting
min_lat <- min(parent_meta_rel$latitude, na.rm=T)
max_lat <- max(parent_meta_rel$latitude, na.rm=T)
min_long <- min(parent_meta_rel$longitude, na.rm=T)
max_long <- max(parent_meta_rel$longitude, na.rm=T)
bbox <- raster::extent(min_long, max_long, min_lat, max_lat)
###########################
# Install required packages if not already

# Download WorldClim bioclimatic variables
clim <- raster::stack("~/Uni/Doctorate/Ch Planting/data/AUS_wc2.1_30s_prec.tif")
annual_prec <- sum(clim)
annual_prec_NSW <- crop(annual_prec, bbox); annual_prec_NSW <- trim(annual_prec_NSW)
annual_prec_Spat <- terra::rast(annual_prec_NSW)

cities <- data.frame(
  city = c("Newcastle", "Port Macquarie", "Coffs Harbour"),
  lat  = c(-32.9283, -31.4349, -30.2962),
  lon  = c(151.7817, 152.9002, 153.1139)
)

Mat_Line_NSW_Plot <- ggplot() +
  #xlim(c(min_long, max_long)) + ylim(c(min_lat, max_lat)) +
  tidyterra::geom_spatraster(data=annual_prec_Spat) +
  geom_point(data = cities, aes(x = lon, y = lat), size = 1, shape = 24, colour = "grey20") +
  geom_text(data = cities, aes(x = lon + 0.1, y = lat, label = city), colour = "grey40", size = 2, hjust = 0) +
  geom_point(data=parent_meta_rel,aes(y=latitude, x=longitude), size = 3, colour="black") +
  geom_point(data=parent_meta_rel,aes(y=latitude, x=longitude, colour = as.numeric(mean_COI))) +
  ggrepel::geom_text_repel(data = parent_meta_rel, aes(x = longitude, y = latitude, label = round(mean_COI,1), colour = as.numeric(mean_COI)), nudge_x = 0.5, box.padding = 0.5, size=3) +
  scale_colour_gradient(low = "chartreuse4", high = "coral2") +
  scale_fill_viridis_c(na.value = NA, option="turbo", direction=-1) +
  xlab(expression(paste("Longitude ("^"o",")"))) + ylab(expression(paste("Latitude ("^"o",")"))) +
  guides(fill=guide_legend(title="Mean annual \nprecipitation (mm)"), colour=guide_legend(title="Mean COI")) +
  theme_bw()

ggsave(Mat_Line_NSW_Plot, filename="/Publication/Figures/Mat_Line_NSW_Plot_SI.jpg", unit='px', width = 1500, height = 2000, limitsize=F)
```


```{r Porometer results data prep}
# Function to prep data...
porom_data_clean <- function(data) {
  colnames <- data[1,]
  data<-data[-c(1:2),]
  colnames(data) <- colnames
  return(data)
}

# Day 15
porometer_results_1 <- read.csv("Porom/Day_15_PorometerMeasurement/K_lowfl_medbth/2025-01-14/K_lowfl_medbth_Karina Guo_2025_01_14_12_26_42_1.csv")
porometer_results_1 <- porom_data_clean(porometer_results_1)

porometer_results_2 <- read.csv("Porom/Day_15_PorometerMeasurement/K_lowfl_medbth/2025-01-15/K_lowfl_medbth_Karina Guo_2025_01_15_08_37_40_1.csv")
porometer_results_2 <- porom_data_clean(porometer_results_2)
porometer_results_2$`Obs#` <- as.numeric(porometer_results_2$`Obs#`) + max(as.numeric(porometer_results_1$`Obs#`))
porometer_results_Day15 <- rbind(porometer_results_1, porometer_results_2)
porometer_results_Day15 <- porometer_results_Day15 %>% 
  dplyr::select(`Obs#`, Time, Date, remark, IID, Drought, gsw) %>% 
  mutate(Set = 'Day15')
  

# Porometer - Day 30
porometer_results_Day30 <- read.csv("Porom/Day_30_PorometerMeasurement/K_lowfl_medbth/2025-01-29/K_lowfl_medbth_Karina Guo_2025_01_29_13_00_09_1.csv")
porometer_results_Day30 <- porom_data_clean(porometer_results_Day30) 

porometer_results_Day30 <- porometer_results_Day30 %>% 
  dplyr::select(`Obs#`, Time, Date, remark, IID, Drought, gsw) %>% 
  mutate(Set = 'Day30')

# Porometer - Day 60
porometer_results_Day60 <- read.csv("Porom/Day_60_PorometerMeasurement/K_lowfl_medbth/2025-02-27/K_lowfl_medbth_Karina Guo_2025_02_27_12_48_54_1.csv")
porometer_results_Day60 <- porom_data_clean(porometer_results_Day60) 

porometer_results_Day60 <- porometer_results_Day60 %>% 
  dplyr::select(`Obs#`, Time, Date, remark, IID, Drought, gsw) %>% 
  mutate(Set = 'Day60')

# Rbind both days
porometer_results <- rbind(porometer_results_Day15, porometer_results_Day30, porometer_results_Day60)
colnames(porometer_results)[(which(names(porometer_results) == "Drought"))] <- "Drought_sev"
porometer_results$Time <- chron::chron(times=porometer_results$Time)
porometer_results$gsw <- as.numeric(porometer_results$gsw)
porometer_results$IID <- as.numeric(porometer_results$IID)
porometer_results$Drought_sev <- factor(porometer_results$Drought_sev, levels = c("Low", "Mid", "High")) # Reorder Sev factor levels

# Joining data with meta
porometer_results_IID <- left_join(porometer_results %>% select(-c('Drought_sev')), IID_block, by = c('IID'))
porometer_results_rmed <- porometer_results_IID %>% filter(gsw<c(-0.05))
porometer_results_IID <- porometer_results_IID %>% filter(gsw>c(-0.05)) #remove impossible results


# Joining local precip + Mat line + porom res 
porometer_results_IID$Drought_len <- as.character(porometer_results_IID$Drought_len)
porometer_results <- porometer_results_IID

porometer_results_median <- porometer_results_IID %>%
  group_by(IID, Set, Mat_Line, Drought_len, Drought_sev) %>%
  summarise(`Obs#` = median(as.numeric(`Obs#`)), median_gsw = median(gsw, na.rm=T), Time=median(Time, na.rm=T)) %>% 
  unique() %>% 
  ungroup()

porometer_results_Mat_Line <- porometer_results_median %>%
  group_by(Set, Mat_Line, Drought_sev, Drought_len) %>%
  summarise(levelmean_gsw = mean(median_gsw)) %>% 
  unique()

porometer_results_Mat_Line_melted <- porometer_results_Mat_Line %>%
  pivot_wider(names_from = "Set", values_from = "levelmean_gsw")
```

```{r Inoculation data prep}
## Inoculation One
# Inoculation data
MR_score_D30 <- read_xlsx("Scoring/MR_Scoring_D30.xlsx", sheet = "Sheet2") %>% mutate(Drought_len="30")
MR_score_D60 <- read_xlsx("Scoring/MR_Scoring_D60.xlsx", sheet = "Sheet2") %>% mutate(Drought_len="60")

  # Joining data with the meta data
MR_score <- rbind(MR_score_D30, MR_score_D60)
MR_score_IID <- left_join(MR_score, IID_block %>% dplyr::select(-Drought_len), by="IID")
MR_score_IID$Drought_sev <- factor(MR_score_IID$Drought_sev, levels = c("Low", "Mid", "High")) # Reorder Sev factor levels
MR_score_IID$Drought_len <- as.character(MR_score_IID$Drought_len)

  # Assign porometer results by representative mat line and drought sev
MR_score_porom <- left_join(MR_score_IID, porometer_results_Mat_Line_melted) %>% 
  filter(!is.na(`COI scores`))

#MR_score_porom$Diff_COI_meanMat <- MR_score_porom$`COI scores`-MR_score_porom$mean_COI

## Inoculation Set Two
# Reinoculation data 
MR_score_D30_inoc2 <- read_xlsx("Scoring/MR_Scoring_D30_Inoc2.xlsx", sheet = "Sheet2")
MR_score_D60_inoc2 <- read_xlsx("Scoring/MR_Scoring_D60_Inoc2.xlsx", sheet = "Sheet2")

  # Joining data with the meta data
MR_score_inoc2 <- rbind(MR_score_D30_inoc2, MR_score_D60_inoc2)
MR_score_inoc2 <- left_join(MR_score_inoc2, IID_block)
MR_score_inoc2$Drought_sev <- factor(MR_score_inoc2$Drought_sev, levels = c("Low", "Mid", "High")) # Reorder Sev factor levels
MR_score_inoc2$Drought_len <- as.character(MR_score_inoc2$Drought_len)

  # Assign porometer results by representative mat line and drought sev
MR_score_inoc2_porom <- left_join(MR_score_inoc2, porometer_results_Mat_Line_melted) %>% 
  filter(!is.na(`COI scores`))

#MR_score_inoc2_porom$Diff_COI <- MR_score_inoc2_porom$`COI scores`-MR_score_inoc2_porom$mean_COI
```


# Script to randomise block
```{r eval=FALSE}
IID_orig <- read.csv("Misc_archive/IID_group_prelim.csv") # Data set for initial IID group allocation

IID_orig_filtBack <- IID_orig %>% filter(!(Type=="Backup"))

final_blocks=NULL

set.seed(12354)

# Assigning blocks with random words
words_assign <- OpenRepGrid::randomWords(1000); words_assign <- words_assign[!duplicated(words_assign)]

count=0 # Counter
for (len in unique(IID_orig_filtBack$Drought_len)) {
  IID_orig_filtBack_rmpres <- IID_orig_filtBack %>% filter (!(IID %in% final_blocks$IID)) # Filter dataset that already is randomised
  IID_orig_lensub <- IID_orig_filtBack_rmpres %>% filter(Drought_len == len)
  
  # Repeat an arbitrary amount of times until dataset is empty
  
  for (iter in 1:50){
    print(paste("Running", iter))
    
    count=count+1 # Counter 
    
    IID_orig_lensub_rmpres <- IID_orig_lensub %>% filter (!(IID %in% final_blocks$IID)) # Filter dataset that already is randomised
    
    print(paste("Remaining IIDs", nrow(IID_orig_lensub_rmpres)))
    
    if (nrow(IID_orig_lensub_rmpres) > 1){
      block <- IID_orig_lensub_rmpres %>% 
        group_by(Drought_sev) %>%
        sample_n(4)
      
      block$block_num <- paste0(count, "_", words_assign[iter])
      
      final_blocks <- rbind(final_blocks, block)
    } else {
      next
    }
  }
}

# Assigning block location

rand_block <- as.data.frame(cbind(block_num = unique(final_blocks$block_num) , block_cell = sample(1:length(unique(final_blocks$block_num))))) # creating a randomised block order
rand_block_ord <- rand_block[order(as.numeric(rand_block$block_cell)),]
rand_block_ord$block_cell <- as.numeric(rand_block_ord$block_cell)

final_blocks_ord <- left_join(final_blocks, rand_block_ord)
final_blocks_ord <- final_blocks_ord[order(final_blocks_ord$block_cell),]



## rand bloc 2 for drought 60
final_blocks_ord <- read.csv ("C:/Users/swirl/OneDrive/Documents/Uni/Doctorate/Ch2 Stressor/block_order_12tray.csv") # Running again for remaining indivs

# Drought 60
final_blocks_ord_filt60 <- final_blocks_ord %>% filter(Drought_len == 60)

rand_block <- as.data.frame(cbind(block_num = unique(final_blocks_ord_filt60$block_num) , block_cell_2 = sample(1:length(unique(final_blocks_ord_filt60$block_num))))) # creating a randomised block order
rand_block_ord <- rand_block[order(as.numeric(rand_block$block_cell_2)),]
rand_block_ord$block_cell_2 <- as.numeric(rand_block_ord$block_cell)

final_blocks_ord <- left_join(final_blocks_ord, rand_block_ord)
final_blocks_ord <- final_blocks_ord[order(final_blocks_ord$block_cell),]
#write.csv(final_blocks_ord, "C:/Users/swirl/OneDrive/Documents/Uni/Doctorate/Ch2 Stressor/block_order_12tray.csv")


######## Which ones to porometer?
## Selecting one for each treatment

library(openxlsx)
final_selection <- read.xlsx("FID Selection & ID & Randomisation_vers1.xlsx", sheet = "real_IID_Block_Cell_12tray")
final_selection <- final_selection[,1:9]

to_porom=NULL

for (i in 1:5) {
  sampled <- final_selection %>%
    group_by(Mat_Line, Drought_sev, Drought_len) %>%
    sample_n(1) %>% # Sample one row without replacement
    mutate(necessity = i) # Assign necessity based on iteration
  
  # Append the sampled rows
  to_porom <- bind_rows(to_porom, sampled)
  
  # Remove sampled rows from the pool for next iteration
  final_selection <- final_selection[!(final_selection$IID %in% to_porom$IID),]
}
```

# Script to determine height of floral foam (AOW RZSM)
```{r}
library(tidyverse)
HistoricalRootZoneSoil <- read.csv ("Publication/Data/AWO_2020-2024_HistoricalRootZoneSoilMoisture_NSW.csv")

colnames(HistoricalRootZoneSoil) <- HistoricalRootZoneSoil[17,]
HistoricalRootZoneSoil <- HistoricalRootZoneSoil[-c(1:17),]

colnames(HistoricalRootZoneSoil) <- c("Date","RZSMPerc", "Month")
HistoricalRootZoneSoil$Date <- as.Date(HistoricalRootZoneSoil$Date, format = "%d/%m/%y")
HistoricalRootZoneSoil$RZSMPerc <- as.numeric(HistoricalRootZoneSoil$RZSMPerc)
HistoricalRootZoneSoil$Month <- as.numeric(HistoricalRootZoneSoil$Month)

season_mapping <- c('1' = "Summer", '2' = "Summer", '3' = "Autumn", '4' = "Autumn", '5' = "Autumn", '6' = "Winter", '7' = "Winter", '8' = "Winter", '9' = "Spring", '10' = "Spring", '11' = "Spring", '12' = "Summer")
HistoricalRootZoneSoil$Season <- season_mapping[HistoricalRootZoneSoil$Month] 



ggplot(HistoricalRootZoneSoil, aes(x=Date, y=RZSMPerc, colour = Season)) +
  geom_hline(yintercept=c(5,10,25,50), colour = "black", linetype="dashed") +
  geom_point() +
  #geom_path() +
  theme_minimal()

# < 50% RZSM
count_track_50 <- NULL
count = 0
for (i in 1:nrow(HistoricalRootZoneSoil)){
  RZSM_row <- HistoricalRootZoneSoil[i,2]
  if (RZSM_row <= 50){
    count = count + 1
  } else {
    count_track_50 <- rbind(count_track_50, count)
    count = 0
  }
}

max(count_track_50) # 138 days

# < 25% RZSM
count_track_25 <- NULL
count = 0
for (i in 1:nrow(HistoricalRootZoneSoil)){
  RZSM_row <- HistoricalRootZoneSoil[i,2]
  if (RZSM_row <= 25){
    count = count + 1
  } else {
    count_track_25 <- rbind(count_track_25, count)
    count = 0
  }
}

max(count_track_25) # 122 days

# < 10% RZSM
count_track_10 <- NULL
count = 0
for (i in 1:nrow(HistoricalRootZoneSoil)){
  RZSM_row <- HistoricalRootZoneSoil[i,2]
  if (RZSM_row <= 10){
    count = count + 1
  } else {
    count_track_10 <- rbind(count_track_10, count)
    count = 0
  }
}

max(count_track_10) # 57 days

# < 5% RZSM
count_track_5 <- NULL
count = 0
for (i in 1:nrow(HistoricalRootZoneSoil)){
  RZSM_row <- HistoricalRootZoneSoil[i,2]
  if (RZSM_row <= 5){
    count = count + 1
  } else {
    count_track_5 <- rbind(count_track_5, count)
    count = 0
  }
}

max(count_track_5) # 33 days

# Plot
count_track_5 <- as.data.frame(count_track_5) %>% mutate(perc = "1_count_track_5")
count_track_10 <- as.data.frame(count_track_10) %>% mutate(perc = "2_count_track_10")
count_track_25 <- as.data.frame(count_track_25) %>% mutate(perc = "3_count_track_25")
count_track_50 <- as.data.frame(count_track_50) %>% mutate(perc = "4_count_track_50")

days_below <- rbind(count_track_5, count_track_10, count_track_25, count_track_50)

ggplot(days_below, aes(x=perc, y=V1)) +
  geom_point() + 
  theme_bw() +
  labs(x="Percentage of RZSM threshold", y="Days in a row below percentage")
  #geom_violin() +

max((count_track_10 %>% dplyr::filter(!(V1==0)))$V1)
```


# Separate variable analysis (Height, weight, porometer, MR scores) & Checks
## Weight
```{r Weight}
# Weight confirmation
height_weight_data_exp <- height_weight_data 

height_weight_data_exp$Drought_sev <- gsub("Mid", "Moderate", height_weight_data_exp$Drought_sev)
height_weight_data_exp$Drought_sev <- factor(height_weight_data_exp$Drought_sev, levels = c("Low", "Moderate", "High"))

height_weight_data_exp$Plot_measure <- ifelse(height_weight_data_exp$Measuring.Purpose == "Pre-experiment", "Before drought", "During drought")

weight_plot <- ggplot(data=height_weight_data_exp, aes(x=Drought_sev, y=Weight..g., fill=Plot_measure)) +
  geom_boxplot() +
  theme_bw() +
  labs (x="Drought severity treatment", y="Weight (g)")
weight_plot

weight_over_time <- ggplot(data=height_weight_data_exp, aes(x=Date, y=Weight..g., colour=Drought_sev, shape = Plot_measure, group = Drought_sev)) +
  stat_smooth(method = 'lm', alpha = 0.1, linewidth = 0.75) +
  geom_vline(xintercept = as.Date("2025-02-02"), linetype = "dashed", color = "firebrick") +
  geom_vline(xintercept = as.Date("2025-03-03"), linetype = "dashed", color = "firebrick") +
  geom_point() +
  theme_bw() +
  scale_colour_manual(labels = c("Low", "Moderate", "High"), values = c("deepskyblue3", "darkolivegreen3", "darkgoldenrod2")) +
  scale_shape_manual(values = c(17, 16)) +
  labs (x="Date", y="Weight (g)", colour = "Drought severity", shape = "Measurement time")
weight_over_time

weight_pre <- lm(Weight..g. ~ Date * Drought_sev, data = (height_weight_data_exp %>% filter ((Measuring.Purpose == "Pre-experiment"))))
anova(weight_pre)

weight_post <- lm(Weight..g. ~ Date * Drought_sev, data = (height_weight_data_exp %>% filter (!(Measuring.Purpose == "Pre-experiment"))))
anova(weight_post)

ggsave(weight_over_time, file="Publication/Figures/weight_plot.jpg", width = 2500, height = 2000, units = "px", limitsize = F)
```

## Analysing Porometer results
Checking if rmed porometer results are represented elsewhere
```{r}
unique_rmed_level <- porometer_results_rmed %>%
  dplyr::select(Drought_sev, Mat_Line) %>% 
  unique() %>% 
  mutate(Level_Identifier=paste0(Mat_Line,Drought_sev))

unique_level <- porometer_results %>% 
  dplyr::select(Drought_sev, Mat_Line) %>% 
  unique() %>% 
  mutate(Level_Identifier=paste0(Mat_Line,Drought_sev))

unique_rmed_level$Level_Identifier %in% unique_level$Level_Identifier  # All represented, we good to go!
```

```{r Porometer results data analysis}
# Does porometer reading correlate to Drought sev?
summary(aov(formula = (Day15 ~ Drought_sev), data = porometer_results_Mat_Line_melted))
summary(aov(formula = (Day30 ~ Drought_sev), data = porometer_results_Mat_Line_melted))
summary(aov(formula = (Day60 ~ Drought_sev), data = porometer_results_Mat_Line_melted))

# Spread of result per IID
# p_all_overall <- ggplot() + 
#   geom_violin(data=porometer_results, aes(x=Drought_sev, y=gsw)) +
#   geom_boxplot(data=porometer_results, aes(x=Drought_sev, y=gsw, colour=as.character(IID), group=IID), position=position_dodge(width=0.75))+
#   facet_wrap(~Set) +
#   theme_bw() +
#   theme(legend.position='none')


# Median of each IID
porometer_results_median <- porometer_results_median %>% 
  filter (!(Set == "Day15"))
porometer_results_median$Set <- gsub("Day", "Day ", porometer_results_median$Set)


sample_sizes <- porometer_results_median %>%
  group_by(Set, Drought_sev) %>%
  summarise(n_count = n(), .groups = 'drop') %>%
  mutate(label_text = paste0("italic(N) == ", n_count),
         y_pos = c(0.2, 0.17, 0.2, 0.25, 0.2, 0.25))
         #y_pos = c(0.61, 0.4, 0.35, 0.2, 0.17, 0.2, 0.25, 0.2, 0.25))

porometer_results_median$Drought_sev <- factor(porometer_results_median$Drought_sev, levels = c("Low", "Mid", "High")) # Reorder Sev factor levels
p_median <- ggplot() + 
  geom_violin(data=porometer_results_median, aes(x=Drought_sev, y=median_gsw)) +
  geom_point(data=porometer_results_median, aes(x=Drought_sev, y=median_gsw, group=Mat_Line, colour=Mat_Line), size=0.5, position = position_dodge(width=0.95)) + 
  facet_wrap(~Set) +
  theme_bw() +
  scale_color_brewer(palette = "Paired") +
  labs(y= "Median gsw measurement", colour="Maternal line", x="Drought severity") +
  scale_x_discrete(labels = c("Low", "Moderate", "High")) +
  geom_text(data = sample_sizes, aes(x = Drought_sev, y = y_pos, label = label_text), color = "grey40", size = 3, parse = TRUE)


#ggsave(p_median, file="Publication/Figures/Day_Droughtsev_mediangsw.jpg", width = 2000, height = 1500, units = "px", limitsize = F)
#   
# tmp <- (porometer_results_median %>% 
#   group_by(Set, Drought_sev, Mat_Line) %>% 
#   summarise(mean_median=mean(median_gsw, na.rm=T))) %>% 
#   group_by(Set, Drought_sev) %>% 
#   mutate(set_sev_mean=mean(mean_median, na.rm=T))

## Maternal line impact
#mquin_meta_filt <- mquin_meta %>% select(FID, Lat, Long) %>% unique(); colnames(mquin_meta_filt)[1] <- "Mat_Line"
#porometer_results_Mat_Line_meta <- left_join(porometer_results, Mat_line_droughttol) 

porometer_results_Mat_Line_meta <- left_join(porometer_results, parent_meta_rel %>% dplyr::select(-c(ID, IID)))  %>% filter(!(Set == "Day15"))
prec_gsw_aov <- aov(formula = (gsw ~ prec*Drought_sev + Drought_len + Set), data = porometer_results_Mat_Line_meta)
summary(prec_gsw_aov)


porometer_results_Mat_Line_meta_median <- left_join(porometer_results_median, parent_meta_rel %>% dplyr::select(-c(ID, IID))) %>% filter(!(Set == "Day15"))

prec_mediangsw_aov_int <- aov(formula = (median_gsw ~ prec*Drought_sev + Drought_len + Set), data = porometer_results_Mat_Line_meta_median)
summary(prec_mediangsw_aov_int)

prec_mediangsw_aov <- aov(formula = (median_gsw ~ prec + Drought_sev + Drought_len + Set), data = porometer_results_Mat_Line_meta_median)
summary(prec_mediangsw_aov)

ggplot(porometer_results_Mat_Line_meta_median, 
             aes(x = (prec), y = median_gsw, color = factor(Drought_sev))) +
  geom_point() +
  stat_summary(fun = mean, geom = "point", size = 3) +
  stat_summary(fun = mean, geom = "line", aes(group = Drought_sev)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2) +
  facet_wrap(~Drought_len, labeller = label_both)

## We measured some individuals more than once at the start and end - how do these compare?
porometer_results_multmeas <- porometer_results %>% 
  filter (Set=="Day60") %>% 
  group_by(IID) %>% 
  mutate (n=n()) %>% 
  filter (n>5)

# ggplot(data=porometer_results_multmeas, aes(x=as.numeric(`Obs#`), y=gsw, colour=Drought_sev)) + 
#   stat_smooth(method='lm', se=F, alpha = 0.8, linetype= 'dashed', linewidth =0.7) +
#   geom_point(position = position_dodge(width=0.75), size=0.75 ) +
#   theme_bw() +
#   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
#   #ylim(min(porometer_results$gsw),max(porometer_results$gsw)) +
#   facet_wrap(~IID)
```

```{r Inoculation score_results of difference between BV}
MR_score_porom %>% group_by(Drought_len, Drought_sev) %>% summarise (n=n()) %>% ungroup()

MR_score_porom$Drought_sev <- factor(MR_score_porom$Drought_sev, levels = c("Low", "Mid", "High")) # Reorder Sev factor levels

MR_Droughtsev_Droughtlen_box_split <- ggplot() +
  geom_boxplot(data = MR_score_porom, aes(x = as.factor(mean_COI), y = (`COI scores`), fill = as.factor(Drought_sev)), linewidth = 0.2) +
  labs(x = "Maternal Line Mean Seedling COI", y = "Post-drought Coefficient of Infection (COI)", fill = "Drought Severity") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_discrete(labels = function(x) sprintf("%.1f", as.numeric(x)))  + # Round to 1 decimal 
  scale_fill_manual(labels = c("Low", "Moderate", "High"), values = c("deepskyblue3", "darkolivegreen3", "darkgoldenrod2")) +
  facet_wrap(~Drought_len, nrow=2)

MR_Droughtsev_Droughtlen_box_split

ggsave(MR_Droughtsev_Droughtlen_box_split, file="Publication/Figures/boxplot_droughtlen_sev.png", width = 8, height = 5)

####
# 
# MR_Droughtsev_Droughtlen_box_split_diff <- ggplot() +
#   geom_boxplot(data = MR_score_porom, aes(x = as.factor(mean_COI), y = Diff_COI_meanMat, fill = as.factor(Drought_sev)), linewidth = 0.2, alpha = 0.5, width=0.7) +
#   #geom_point(data = MR_score_porom, aes(x = as.factor(mean_COI), y = Diff_COI_meanMat, colour = as.factor(Drought_sev)), alpha = 0.5, ,position = position_jitterdodge( seed=1, dodge.width = 0.75)) +
#   labs(x = "Maternal Line Mean Seedling COI", y = "Post-drought COI - Maternal Line Mean Seedling COI", fill = "Drought Severity", colour = "Drought Severity") +
#   theme_bw() +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
#   scale_x_discrete(labels = function(x) sprintf("%.1f", as.numeric(x)))  + # Round to 1 decimal 
#   scale_fill_manual(labels = c("Low", "Moderate", "High"), values = c("skyblue2", "darkolivegreen2", "goldenrod1")) +
#   scale_colour_manual(labels = c("Low", "Moderate", "High"), values = c("deepskyblue4", "darkolivegreen4", "darkgoldenrod3")) +
#   facet_wrap(~Drought_len, nrow=2)
# 
# MR_Droughtsev_Droughtlen_box_split_diff
```


# Seedling GT Data

## Process data
```{r}
library(tidyverse); library(readxl); library(logger)

#######################################
# Genomic prediction
source("C:/Users/swirl/OneDrive/Documents/Uni/Doctorate/Samples/Genotyping/Code/GP_functs.R")

base_dir="~/Uni/Doctorate/Samples/Genotyping/Report-DMela25-10753/Report-DMela25-10753/"
  
gt_datafile=paste0(base_dir, "Report-DMela25-10753_trainingconcat_sort.hapmap.hmp.txt")
input_phenodatafile="~/RBGSyd_Technical Officer/MQuin/Processing Meta/mq_phenotypes.csv"
out_dir=paste0(base_dir, "GP_Out_rmclimsnps/Nohomo/")
log_file=paste0(base_dir, "log_file_withclim_nohomo.txt")

run_genomic_prediction(gt_datafile = gt_datafile, input_phenodatafile = input_phenodatafile, trans_COI=TRUE, rm_clim_snp=TRUE, homo_site_filt=1, miss_site_filt=0.6, homo_samp_filt=1, miss_samp_filt=0.9, out_dir = out_dir, log_file = log_file) 
```


## Joining against scoring
```{r, warning = FALSE}
base_dir="~/Uni/Doctorate/Samples/Genotyping/Report-DMela25-10753/Report-DMela25-10753/"

prediction_gt <- read.csv(paste0(base_dir, "GP_Out_rmclimsnps/Nohomo/prediction_gt.csv")) %>% dplyr::select(c(Taxa, PEV, Prediction))

colnames(prediction_gt)[1] <- "NSWID"
sample_meta_drought <- read_xlsx("~/Uni/Doctorate/Samples/All_samples.xlsx") %>% filter(!is.na(NSWID) & `Genotyping Purpose`=="Drought_experiment") %>% dplyr::select(NSWID, Drought_Exp_ID)
prediction_gt_droughtexp <- inner_join(prediction_gt, sample_meta_drought, by="NSWID")

## Removing low reliability predictions
prediction_gt_droughtexp_poorsamp <- prediction_gt_droughtexp %>% filter(PEV > 0.5)
prediction_gt_droughtexp <- prediction_gt_droughtexp %>% filter(PEV < 0.5)
cat(nrow(prediction_gt_droughtexp_poorsamp), "samples removed by PEV \n")
colnames(prediction_gt_droughtexp)[4] <- "IID"; prediction_gt_droughtexp$IID <- as.numeric(prediction_gt_droughtexp$IID )
colnames(prediction_gt_droughtexp)[3] <- "Prediction_IID"

## Merging with scoring data
MR_score_GT_pred <- left_join(prediction_gt_droughtexp, MR_score_IID, by="IID")
MR_score_pred_GT <- left_join(MR_score_IID, prediction_gt_droughtexp, by="IID")

### Total genotyoped
nrow(MR_score_GT_pred %>% filter (!is.na(Prediction_IID)))
```
## Basic analyses
We're not expecting this to be directly correlated, as these should show augmented drought values rather than optimal conditions
```{r}
MR_score_pred_GT$Drought_sev <- gsub("Mid", "Moderate", MR_score_pred_GT$Drought_sev)
MR_score_pred_GT$Drought_sev <- factor(MR_score_pred_GT$Drought_sev, levels = c("Low", "Moderate", "High"))

plot_facet_sev <- ggplot(data=MR_score_pred_GT, aes(x=Prediction_IID, y=(`COI scores`)^0.25, colour = as.factor(Drought_sev), fill = as.factor(Drought_sev))) +
  geom_point(alpha=0.5) +
  geom_point(colour="grey70", alpha=0.3) +
  geom_smooth(method = 'lm', se=T, alpha=0.1) +
  geom_smooth(method = 'lm', se=F) +
  theme_bw() +
  geom_abline(intercept=0, slope=1, colour="grey", linetype="dashed") +
  labs(y=expression((paste("Post-drought COI score"))^0.25), x="Genomically Estimated Breeding Value for COI", colour="Drought severity") +
  scale_colour_manual(values = c("deepskyblue3", "darkolivegreen3", "darkgoldenrod2")) +
  scale_fill_manual(values = c("deepskyblue3", "darkolivegreen3", "darkgoldenrod2")) +
  theme(legend.position = "bottom") +
  guides(fill="none")

plot_facet_sev_marg <- ggExtra::ggMarginal(plot_facet_sev, type="density", groupColour = TRUE, groupFill = TRUE, alpha=0.3) 
plot_facet_sev_marg

## Add images of susceptible and resistant

#ggsave(plot_facet_sev_marg, filename="Publication/Figures/GT_GEBV_Droughtsev.jpg", unit='px', width = 2500, height = 2000, limitsize=F)

plot_facet_mat <- ggplot(MR_score_GT_pred, aes(x=Prediction_IID, y=(`COI scores`)^0.25, colour = Mat_Line)) +
  geom_point() +
  geom_abline(intercept=0, slope=1, colour="grey", linetype="dashed") +
  geom_smooth(method = 'lm') +
  theme_bw() +
  labs(y=expression((COI~scores)^0.25)) +
  facet_wrap(~Mat_Line * Drought_sev)

#plot_facet_mat
```

## Departure from expected - 
A) is it  higher at the two extremes?
B) is it correlated by maternal line?
C) What about A) for Day 30 or 60

```{r}
MR_score_GT_pred$Difference_pred_IID <- MR_score_GT_pred$Prediction_IID - (MR_score_GT_pred$`COI scores`)^0.25
MR_score_GT_pred_rmNA <- MR_score_GT_pred %>% 
  filter(
    !is.na(Prediction_IID), 
    !is.na(Drought_sev), 
    !is.na(Drought_len)
  )
# A) is it  higher at the two extremes?
  # Visualisation
plot_diff <- ggplot() +
  geom_violin(data=MR_score_GT_pred_rmNA, aes(x=Drought_sev, y=Difference_pred_IID), alpha=0, colour="grey60") +
  stat_summary(data=MR_score_GT_pred_rmNA, aes(x=Drought_sev, y=Difference_pred_IID), fun.y = mean, geom = "point", size = 20, colour = "grey10", shape = 3) +
  geom_boxplot(data=MR_score_GT_pred_rmNA, aes(x=Drought_sev, y=Difference_pred_IID, color = Mat_Line), alpha=0.5) +
  #geom_point(data=MR_score_GT_pred_rmNA, aes(x=Drought_sev, y=Difference_pred_IID, color = Mat_Line), alpha=0.5, position = position_jitterdodge(jitter.width = 0.05, dodge.width = 0.8)) +
  facet_wrap(~Drought_len) +
  scale_colour_manual( name = "Maternal line", values = scales::alpha(RColorBrewer::brewer.pal(10, "Paired"), 0.8)) +
  theme_bw() +
  scale_x_discrete(labels = c("Low", "Moderate", "High")) +
  labs(x="Drought severity", y="Coefficient of infection difference")
  
#ggsave(plot_diff, file="Publication/Figures/Mat_Line_Droughtsev_COIDiff.jpg", width=2500, height=1750, limitsize=F, units="px")


# B) is it correlated by maternal line?
  # Visualisation
plot_diff_facet <- ggplot() +
  geom_boxplot(data=MR_score_GT_pred_rmNA, aes(x=Drought_sev, y=Difference_pred_IID)) +
  geom_jitter(data=MR_score_GT_pred_rmNA, aes(x=Drought_sev, y=Difference_pred_IID), alpha=0.5) +
  theme_bw() +
  facet_wrap(~Mat_Line)

plot_diff_facet

parent_meta_rel_sub <- parent_meta_rel %>% dplyr::select(Mat_Line, latitude, longitude, prec) 
MR_score_GT_pred_rmNA_meta <- left_join(MR_score_GT_pred_rmNA, parent_meta_rel_sub)

  # Statistics
Departure_A <- aov(data = MR_score_GT_pred_rmNA_meta, formula = Difference_pred_IID ~ Mat_Line*Drought_sev + Drought_len);Departure_A
summary(Departure_A)
# TukeyHSD(Departure_A, "Drought_sev", ordered = TRUE)
# TukeyHSD(Departure_A, "Drought_len", ordered = TRUE)
# TukeyHSD(Departure_A, "Mat_Line", ordered = TRUE)
# ggpubr::ggqqplot(residuals(Departure_A))

# MAP
Departure_A_prec_int <- aov(data = MR_score_GT_pred_rmNA_meta, formula = Difference_pred_IID ~ prec * Drought_sev + Drought_len);Departure_A_prec_int
summary(Departure_A_prec_int)

Departure_A_prec <- aov(data = MR_score_GT_pred_rmNA_meta, formula = Difference_pred_IID ~ prec + Drought_sev + Drought_len);Departure_A_prec
summary(Departure_A_prec)

Departure_A_raw_prec_int <- aov(data = MR_score_GT_pred_rmNA_meta, formula = `COI scores`^0.25 ~ Drought_sev * prec + Drought_len); Departure_A_raw_prec_int
summary(Departure_A_raw_prec_int)

Departure_A_raw_prec <- aov(data = MR_score_GT_pred_rmNA_meta, formula = `COI scores`^0.25 ~ Drought_sev + prec + Drought_len); Departure_A_raw_prec
summary(Departure_A_raw_prec)

summary(prec_mediangsw_aov)

TukeyHSD(Departure_A_raw_prec, "Drought_sev", ordered = TRUE)
TukeyHSD(Departure_A_raw_prec, "Drought_len", ordered = TRUE)

# C) What about A) for Day 30 or 60
plot_droughtlen <- ggplot() +
  geom_boxplot(data=MR_score_GT_pred_rmNA, aes(x=Drought_sev, y=Difference_pred_IID, fill = Drought_len), alpha=0.1, position = position_dodge(width = 0.8)) +
  geom_point(data=MR_score_GT_pred_rmNA, aes(x=Drought_sev, y=Difference_pred_IID, color = Mat_Line, group = Drought_len), alpha=0.5, position = position_jitterdodge(jitter.width = 0.05, dodge.width = 0.7)) +
  theme_bw()

plot_droughtlen

Departure_B <- aov(data = MR_score_GT_pred_rmNA, formula = Difference_pred_IID ~ Drought_sev + Drought_len + Mat_Line)
summary(Departure_B); TukeyHSD(Departure_B, ordered = TRUE)

Departure_B_raw <- aov(data = MR_score_GT_pred_rmNA, formula = `COI scores`^0.25 ~ Drought_sev + Drought_len + Mat_Line)
summary(Departure_B_raw); TukeyHSD(Departure_B_raw, ordered = TRUE)

(TukeyHSD(Departure_B_raw, ordered = TRUE)$Mat_Line)[3:17,4] < 0.05

# omitted_rows <- na.action(Departure_B)
# excluded_data <- MR_score_GT_pred_rmNA_meta[omitted_rows, ]


```

```{r}
# Residuals 

Excess_res_lat_mat <- aov(data = MR_score_GT_pred_rmNA_meta, formula = Difference_pred_IID ~ Drought_sev + Drought_len + prec + Mat_Line)
Excess_res_lat <- aov(data = MR_score_GT_pred_rmNA_meta, formula = Difference_pred_IID ~ Drought_sev + Drought_len + prec)

MR_score_GT_pred_rmNA_meta_res <- MR_score_GT_pred_rmNA_meta %>% 
  filter(!is.na(Difference_pred_IID))

MR_score_GT_pred_rmNA_meta_res$residuals_latmat <- residuals(Excess_res_lat_mat)
MR_score_GT_pred_rmNA_meta_res$residuals_lat <- residuals(Excess_res_lat)

ggplot(MR_score_GT_pred_rmNA_meta_res, aes(x = prec, y = residuals_latmat)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "blue") +
  labs(title = "LatMat residual")

ggplot(MR_score_GT_pred_rmNA_meta_res, aes(x = prec, y = residuals_lat)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "blue") +
  labs(title = "Lat residual")

```


# Reinoculation
```{r Reinoculation}
MR_score_D30 <- read_xlsx("Scoring/MR_Scoring_D30.xlsx", sheet = "Sheet2") %>% mutate(Drought_len="30")
MR_score_D60 <- read_xlsx("Scoring/MR_Scoring_D60.xlsx", sheet = "Sheet2") %>% mutate(Drought_len="60")
MR_score_D30_inoc2 <- read_xlsx("Scoring/MR_Scoring_D30_Inoc2.xlsx", sheet = "Sheet2")
MR_score_D60_inoc2 <- read_xlsx("Scoring/MR_Scoring_D60_Inoc2.xlsx", sheet = "Sheet2")

# Checking escapees
MR_score_D30_0 <- MR_score_D30 %>% 
  filter (`COI scores` == 0)
MR_score_D30_inoc2_0 <- MR_score_D30_inoc2 %>% 
  filter (IID %in% MR_score_D30_0$IID)

prop.table(table(ifelse(MR_score_D30_inoc2_0$`COI scores` == 0, "HR", "Escaped")))

MR_score_D60_0 <- MR_score_D60 %>% 
  filter (`COI scores` == 0)
MR_score_D60_inoc2_0 <- MR_score_D60_inoc2 %>% 
  filter (IID %in% MR_score_D60_0$IID)

prop.table(table(ifelse(MR_score_D60_inoc2_0$`COI scores` == 0, "HR", "Escaped")))

# Directional shift
# 1. How many changed COI in the second set of inoculation
MR_score_comp_1 <- MR_score %>% 
  dplyr::select(IID,`COI scores`, Drought_len) %>%
  unique()
colnames(MR_score_comp_1) <- c('IID', 'COI_Scores_Inoc1', 'Drought_len')
MR_score_comp_1$Drought_len <- as.numeric(MR_score_comp_1$Drought_len)

MR_score_comp_2 <- MR_score_inoc2 %>% 
  dplyr::select(IID,`COI scores`, Drought_len) %>%
  unique()
colnames(MR_score_comp_2) <- c('IID', 'COI_Scores_Inoc2', 'Drought_len')
MR_score_comp_2$Drought_len <- as.numeric(MR_score_comp_2$Drought_len)

MR_score_comp <- left_join(MR_score_comp_1, MR_score_comp_2) 

MR_score_comp_rmna <- MR_score_comp %>%
  filter(!is.na(COI_Scores_Inoc2))

MR_score_comp_rmna_rm0inoc1 <- MR_score_comp_rmna %>% 
  filter(!(COI_Scores_Inoc1 == 0))

MR_score_comp_rmna_rm0inoc1$Difference <- MR_score_comp_rmna_rm0inoc1$COI_Scores_Inoc2 - MR_score_comp_rmna_rm0inoc1$COI_Scores_Inoc1

# Split by drought sev - D30
MR_score_reinoc_comp_meta <- left_join(MR_score_comp_rmna_rm0inoc1, IID_block) %>% 
  unique()

MR_score_reinoc_comp_meta$Drought_sev <- gsub("Mid", "Moderate",MR_score_reinoc_comp_meta$Drought_sev)
MR_score_reinoc_comp_meta$Drought_sev <- factor(MR_score_reinoc_comp_meta$Drought_sev, levels = c("Low", "Moderate", "High"))

# Keep escapees for SI
MR_score_comp_rmna_rm0inoc1_keepescape <- MR_score_comp_rmna 

MR_score_comp_rmna_rm0inoc1_keepescape$Difference <- MR_score_comp_rmna_rm0inoc1_keepescape$COI_Scores_Inoc2 - MR_score_comp_rmna_rm0inoc1_keepescape$COI_Scores_Inoc1

# Split by drought sev - D30
MR_score_reinoc_comp_meta_keepescape <- left_join(MR_score_comp_rmna_rm0inoc1_keepescape, IID_block) %>% 
  unique()

MR_score_reinoc_comp_meta_keepescape$Drought_sev <- gsub("Mid", "Moderate",MR_score_reinoc_comp_meta_keepescape$Drought_sev)
MR_score_reinoc_comp_meta_keepescape$Drought_sev <- factor(MR_score_reinoc_comp_meta_keepescape$Drought_sev, levels = c("Low", "Moderate", "High"))

  # Visualisation
split_shift_lm <- ggplot() +
  #geom_point(data = MR_score_reinoc_comp_meta_keepescape, aes(x=COI_Scores_Inoc1^0.25, y=COI_Scores_Inoc2^0.25, colour = Difference), alpha=0.5) +
  geom_point(data = MR_score_reinoc_comp_meta, aes(x=COI_Scores_Inoc1^0.25, y=COI_Scores_Inoc2^0.25, colour = Difference), alpha=0.5) +
  theme_bw () + 
  geom_abline(intercept=0, slope=1, colour="grey", linetype="dashed") +
  facet_wrap (~Drought_len * Drought_sev) +
  labs (x=expression(paste("Droughted score (COI)")^0.25), y=expression(paste("Water replete score (COI)")^0.25), colour = paste("Difference \n(untransformed COI)"))

split_shift_lm

#ggsave(split_shift_lm, filename="Publication/Figures/Droughted_v_Waterprelete.jpg", unit='px', width = 2000, height = 1500, limitsize=F)
#ggsave(split_shift_lm, filename="Publication/Figures/Droughted_v_Waterprelete_SI.jpg", unit='px', width = 2000, height = 1500, limitsize=F)

## Paired t-test
ShiftInoc_paired <- t.test(MR_score_reinoc_comp_meta$COI_Scores_Inoc2^0.25, MR_score_reinoc_comp_meta$COI_Scores_Inoc1^0.25, paired=TRUE)
ShiftInoc_paired
# 
# ShiftInoc <- aov(data = MR_score_reinoc_comp_meta, formula = Difference ~ Drought_sev + Drought_len)
# summary(ShiftInoc)
```

# Comparing against predicted resistance
```{r}
MR_score_GT_pred$IID <- as.numeric(MR_score_GT_pred$IID); MR_score_GT_pred$Drought_len <- as.numeric(MR_score_GT_pred$Drought_len)
MR_score_reinoc_comp_pred <- left_join(MR_score_comp_rmna_rm0inoc1, MR_score_GT_pred) %>% 
  filter(!is.na(Prediction_IID))
MR_score_reinoc_comp_pred$Drought_sev <- gsub("Mid", "Moderate",MR_score_reinoc_comp_pred$Drought_sev)
MR_score_reinoc_comp_pred$Drought_sev <- factor(MR_score_reinoc_comp_pred$Drought_sev, levels = c("Low", "Moderate", "High"))


####
# No escapees
MR_score_GT_pred$IID <- as.numeric(MR_score_GT_pred$IID); MR_score_GT_pred$Drought_len <- as.numeric(MR_score_GT_pred$Drought_len)
MR_score_reinoc_comp_pred_rmescape <- left_join(MR_score_comp_rmna, MR_score_GT_pred) %>% 
  filter(!is.na(Prediction_IID)) %>% 
  filter (!(COI_Scores_Inoc1 == 0))
MR_score_reinoc_comp_pred_rmescape$Drought_sev <- gsub("Mid", "Moderate",MR_score_reinoc_comp_pred_rmescape$Drought_sev)
MR_score_reinoc_comp_pred_rmescape$Drought_sev <- factor(MR_score_reinoc_comp_pred_rmescape$Drought_sev, levels = c("Low", "Moderate", "High"))


plot_COI_Scores_Inoc1 <- 
  #ggplot(data = MR_score_reinoc_comp_pred, aes(x=Prediction_IID, y=COI_Scores_Inoc1^0.25)) +
  ggplot(data = MR_score_reinoc_comp_pred_rmescape, aes(x=Prediction_IID, y=COI_Scores_Inoc1^0.25)) +
  theme_bw () + 
  geom_smooth (method = 'lm', colour="goldenrod", linewidth=0.7, se=F) +
  scale_linetype_manual(values = c("Low" = "solid", "Moderate" = "dashed", "High" = "solid"), 
                        guide = "none") +
  geom_abline(intercept=0, slope=1, colour="grey", linetype="dashed") +
  geom_point(alpha=0.5, fill ="black", colour = "transparent", shape = 21) +
  facet_wrap (~Drought_len * Drought_sev) +
  labs (x="Genomically Estimated Breeding Value for COI", y=expression("Droughted score (COI)"^0.25))

plot_COI_Scores_Inoc2 <- 
  ggplot(data = MR_score_reinoc_comp_pred_rmescape, aes(x=Prediction_IID, y=COI_Scores_Inoc2^0.25)) +
  #ggplot(data = MR_score_reinoc_comp_pred, aes(x=Prediction_IID, y=COI_Scores_Inoc2^0.25)) +
  theme_bw () + 
  geom_smooth (method = 'lm', colour="steelblue3", linewidth=0.7, se=F) +
  geom_abline(intercept=0, slope=1, colour="grey", linetype="dashed") +
  geom_point(alpha=0.5,fill ="black", colour = "transparent", shape = 21) +
  facet_wrap (~Drought_len * Drought_sev) +
  labs (x="Genomically Estimated Breeding Value for COI", y=expression("Water replete score (COI)"^0.25)) 
plot_COI_Scores_Inoc1
plot_COI_Scores_Inoc2
#ggsave(plot_COI_Scores_Inoc1, , filename="Publication/Figures/GT_GEBV_Reinoc1.jpg", unit='px', width = 1500, height = 1750, limitsize=F)
#ggsave(plot_COI_Scores_Inoc2, filename="Publication/Figures/GT_GEBV_Reinoc2.jpg", unit='px', width = 1500, height = 1750, limitsize=F)
      
model_1 <- lm(data = MR_score_reinoc_comp_pred_rmescape, formula = COI_Scores_Inoc1^0.25 ~ Prediction_IID * Drought_sev + Drought_len)
model_2 <- lm(data = MR_score_reinoc_comp_pred_rmescape, formula = COI_Scores_Inoc2^0.25 ~ Prediction_IID * Drought_sev + Drought_len)

emmeans::emtrends(model_1, ~ Drought_sev * Drought_len, var = "Prediction_IID")
emmeans::emtrends(model_2, ~ Drought_sev * Drought_len, var = "Prediction_IID")

####

# Keep escapees for SI
MR_score_GT_pred$IID <- as.numeric(MR_score_GT_pred$IID); MR_score_GT_pred$Drought_len <- as.numeric(MR_score_GT_pred$Drought_len)
MR_score_reinoc_comp_pred_keepescape <- left_join(MR_score_comp_rmna, MR_score_GT_pred) %>% 
  filter(!is.na(Prediction_IID))
MR_score_reinoc_comp_pred_keepescape$Drought_sev <- gsub("Mid", "Moderate",MR_score_reinoc_comp_pred_keepescape$Drought_sev)
MR_score_reinoc_comp_pred_keepescape$Drought_sev <- factor(MR_score_reinoc_comp_pred_keepescape$Drought_sev, levels = c("Low", "Moderate", "High"))


plot_COI_Scores_Inoc1 <- 
  ggplot(data = MR_score_reinoc_comp_pred_keepescape, aes(x=Prediction_IID, y=COI_Scores_Inoc1^0.25)) +
  theme_bw() + 
  geom_smooth(aes(linetype = Drought_sev), method = 'lm', colour="goldenrod", linewidth=0.7, se=F) +
  scale_linetype_manual(values = c("Low" = "solid", "Moderate" = "solid", "High" = "dashed"), 
                        guide = "none") +
  geom_abline(intercept=0, slope=1, colour="grey", linetype="dashed") +
  geom_point(alpha=0.5, fill ="black", colour = "transparent", shape = 21) +
  facet_wrap(~Drought_len * Drought_sev) +
  labs(x="Genomically Estimated Breeding Value for COI", y=expression("Droughted score (COI)"^0.25))

plot_COI_Scores_Inoc2 <- 
  ggplot(data = MR_score_reinoc_comp_pred_keepescape, aes(x=Prediction_IID, y=COI_Scores_Inoc2^0.25)) +
  #ggplot(data = MR_score_reinoc_comp_pred, aes(x=Prediction_IID, y=COI_Scores_Inoc2^0.25)) +
  theme_bw () + 
  geom_smooth (method = 'lm', colour="steelblue3", linewidth=0.7, se=F) +
  geom_abline(intercept=0, slope=1, colour="grey", linetype="dashed") +
  geom_point(alpha=0.5,fill ="black", colour = "transparent", shape = 21) +
  facet_wrap (~Drought_len * Drought_sev) +
  labs (x="Genomically Estimated Breeding Value for COI", y=expression("Water replete score (COI)"^0.25)) 
plot_COI_Scores_Inoc1
plot_COI_Scores_Inoc2
#ggsave(plot_COI_Scores_Inoc1, , filename="Publication/Figures/GT_GEBV_Reinoc1_SI.jpg", unit='px', width = 1500, height = 1750, limitsize=F)
#ggsave(plot_COI_Scores_Inoc2, filename="Publication/Figures/GT_GEBV_Reinoc2_SI.jpg", unit='px', width = 1500, height = 1750, limitsize=F)

      
model_1 <- lm(data = MR_score_reinoc_comp_pred_keepescape, formula = COI_Scores_Inoc1^0.25 ~ Prediction_IID * Drought_sev + Drought_len)
model_2 <- lm(data = MR_score_reinoc_comp_pred_keepescape, formula = COI_Scores_Inoc2^0.25 ~ Prediction_IID * Drought_sev + Drought_len)

emmeans::emtrends(model_1, ~ Drought_sev * Drought_len, var = "Prediction_IID")
emmeans::emtrends(model_2, ~ Drought_sev * Drought_len, var = "Prediction_IID")

      

####
 
summary(aov(formula = (COI_Scores_Inoc1^0.25) ~ Prediction_IID + Drought_len + Drought_sev, data=MR_score_reinoc_comp_pred))

# Reshape into long format
long_dat <- MR_score_reinoc_comp_pred %>%
  transmute(
    Prediction_IID,
    Inoc1 = COI_Scores_Inoc1^0.25,
    Inoc2 = COI_Scores_Inoc2^0.25
  ) %>%
  pivot_longer(cols = c(Inoc1, Inoc2),
               names_to = "Inoc",
               values_to = "COI_Scores")

# Fit model with interaction
lm_res <- lm(COI_Scores ~ Prediction_IID * Inoc, data = long_dat)

# Check if slopes differ between Inoc1 and Inoc2
anova(lm_res)
summary(lm_res)


#### Correlation between water replete - Drought difference vs GEBV?
MR_score_reinoc_comp_pred$COI_Scores_Inoc21_forthtrans <- MR_score_reinoc_comp_pred$COI_Scores_Inoc2^0.25-MR_score_reinoc_comp_pred$COI_Scores_Inoc1^0.25

ggplot(data = MR_score_reinoc_comp_pred, aes(x=Prediction_IID, y=COI_Scores_Inoc21_forthtrans), alpha=0.5) +
  theme_bw () + 
  geom_smooth (method = 'lm', colour="chartreuse4", linetype="dashed", linewidth=0.7, se=F) +
  geom_abline(intercept=0, slope=1, colour="grey", linetype="dashed") +
  geom_point(alpha=0.5,fill ="black", colour = "transparent", shape = 21) +
  #facet_wrap (~Drought_len * Drought_sev) +
  labs (x="Genomically Estimated Breeding Value for COI", y="Water-replete score - Droughted score")
```

# Plot hypotheses
```{r}
hyp_data <- data.frame(x=(runif(n=500, min=0, max=3)), y=(runif(n=500, min=0, max=3)))

hypothesis_plot <- ggplot(data = hyp_data, aes(x, y)) +
  theme_bw () + 
  geom_abline(intercept=0, slope=1, colour="grey", linetype="dashed") +
  #geom_point(alpha=0.3,fill ="black", colour = "transparent", shape = 21) +
  #facet_wrap (~Drought_len * Drought_sev) +
  labs (x="Baseline Coefficient of Infection score using genomic prediction (GEBV)", y="Coefficient of Infection score post-drought treatment")

ggsave(hypothesis_plot, filename="/Publication/Figures/Hypothesis_plot.jpg", unit='px', width = 1500, height = 1500, limitsize=F)
```


